Artificial Ecosystem Selection: Methane Oxidation

### 2020-10-22 R/process.Rmd
recover processing rmd file to look at fluxes

```{r eval=FALSE, include=FALSE}
deviance <-
  fluxes %>% 
  select(treat, passage, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(mean = mean(estimate), .groups = "drop") %>% 
  pivot_wider(names_from = treat, values_from = mean) %>% 
  mutate(deviance = p - n)

se <- function(x, na.rm = FALSE){
  sqrt(var(if (is.vector(x) || is.factor(x)) x else as.double(x), 
    na.rm = na.rm)) / sqrt(length(if (na.rm == TRUE) na.omit(x) else x))
}

sed <- function(p, n, na.rm = FALSE) {
  sqrt(se(p, na.rm = na.rm) ^ 2 + se(n, na.rm = na.rm)^ 2)
}

deviance$se <-
  fluxes %>% 
  select(treat, passage, estimate) %>% 
  mutate(x = 1:nrow(fluxes)) %>% 
  pivot_wider(names_from = treat, values_from = estimate) %>% 
  group_by(passage) %>% 
  summarize(sed = sed(p, n, na.rm = TRUE), .groups = "drop") %>% 
  pull(sed)

fit <- rma(deviance ~ passage, se^2, method="FE", data = deviance)

ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
  geom_pointrange(color = 'darkorange2') + 
  geom_abline(intercept = fit$beta[1], slope = fit$beta[2], color = 'blue')

rma_output <- data.frame(
Estimate = fit$beta,
se = fit$se,
z.value = fit$zval,
p.value = fit$pval,
upper.ci = fit$ci.lb,
lower.ci = fit$ci.ub
)
rma_output %>% 
  kable() %>% 
  kable_styling()

write.csv(deviance, 
```

@@

```{r heritability, message=FALSE, warning=FALSE}
selected_jars <- read_csv('../Data/selected.csv')

parental <- 
  fluxes[paste0(fluxes$passage, fluxes$jar) %in% paste0(selected_jars$passage, selected_jars$jar), ] %>% 
  ungroup() %>% 
  select(treat, rep, estimate, passage)

# Mean parental (selected) flux for each passage
parental <-
  parental %>% 
  group_by(passage, treat) %>% 
  summarize(parental = mean(estimate), .groups = "drop")

offspring <- 
  fluxes %>% 
  ungroup() %>%
  select(treat, rep, estimate, passage)

# Mean offspring flux for each passage
offspring <-
  offspring %>% 
  group_by(passage, treat) %>%
  summarize(offspring = mean(estimate), .groups = "drop") %>%
  filter(passage != 1) %>% 
  mutate(passage = passage - 1)
```

```{r}

###################################################
# This is the nest place you stopped
# You need to calculate the slope of midoffspring
# on midparent for the Positive treatments to get R
# and then calculate the selection differential
# as above S. S doesn't make sense for the neutral
# treatment.
###################################################

# offspring <-
#   fluxes %>%
#   ungroup() %>%
#   select(estimate, passage, treat) %>%
#   mutate(offspring = rank(estimate))

heritability <- left_join(parental, offspring, by = c("passage", "treat"))
heritability$passage <- factor(heritability$passage)
heritability$treat <- factor(heritability$treat)
heritability <-
  filter(heritability, passage != 5)
# heritability$parental <- rank(heritability$parental)

@@

ggplot(heritability, aes(x = parental, y = offspring)) +
  geom_point(aes(color = treat)) +
  stat_smooth(method = 'lm') + 
  labs(x = "Parental Mean Flux (-k)", y = "Offspring Flux (-k)") +
  scale_color_discrete(name = "Selection\nTreatment", labels = c('Neutral', 'Positive')) +
  theme_bw()

# ggsave('heritability.pdf', width = 4, height = 3)
# ggsave('heritability.png', width = 4, height = 3)

```
```{r}
# This chunk calculates the slope of mid-offspring on mid-parent for all generations and all treatments

fluxes
```

```{r breeders}

h2 <- function(R, S) {
  R / S
}

# R = Response to selection, difference in mean between offspring of selected parents and all of the parents

# S = Selection differental, difference in mean between selected parents and all of the parents

parental_all <-
  fluxes %>% 
  select(passage, treat, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(all = mean(estimate), .groups = "drop")

parental_selected <- parental

selection_diff <- 
  left_join(parental_selected, parental_all) %>% 
  rename(selected = parental) %>% 
  mutate(S = selected - all)

heritability_ns <- 
  parental_all %>% 
  left_join(parental_selected) %>% 
  left_join(offspring) %>% 
  filter(passage != 5)

heritability_only_p <- 
  heritability_ns %>% 
  filter(treat == 'p')

R <- tidy(lm(offspring ~ all, data = heritability_only_p))[[2, 2]]
S <- tidy(lm(parental ~ all, data = heritability_only_p))[[2, 2]]
h2(R, S)

# Generational R, S, global h2, 0.82
heritability_only_p %>% 
  mutate(R = offspring - all,
         S = parental - all) %>% 
  summarize(R = mean(R),
            S = mean(S)) %>%
  mutate(h2 = h2(R, S))

# Generational R, S, h2, 0.73
heritability_only_p %>% 
  #filter(passage != 2) %>% 
  mutate(R = offspring - all,
         S = parental - all,
         h2 = h2(R, S)) %>% 
  summarize(se = se(h2),
            h2 = mean(h2))

# Global R, S, h2, 0.82
heritability_only_p %>% 
  summarize(offspring = mean(offspring),
         all = mean(all),
         parental = mean(parental)) %>% 
  mutate(R = offspring - all,
         S = parental - all,
         h2 = h2(R, S))
heritability_ns %>% 
  pivot_longer(all:offspring, names_to = "group", values_to = "mean") %>% 
ggplot(., aes(x = passage, y = mean, color = group)) +
  facet_wrap(~ treat) +
  geom_point()
```

>>>>

```{r deviance, eval=FALSE, include=FALSE}
# Calculate deviance as p - n
deviance <-
  fluxes %>% 
  select(treat, passage, estimate) %>% 
  group_by(passage, treat) %>% 
  summarize(mean = mean(estimate), .groups = "drop") %>% 
  pivot_wider(names_from = treat, values_from = mean) %>% 
  mutate(deviance = p - n)

# Create se function that can handle NAs
se <- function(x, na.rm = FALSE){
  sqrt(var(if (is.vector(x) || is.factor(x)) x else as.double(x), 
    na.rm = na.rm)) / sqrt(length(if (na.rm == TRUE) na.omit(x) else x))
}

# Standard error of the difference function
sed <- function(p, n, na.rm = FALSE) {
  sqrt(se(p, na.rm = na.rm) ^ 2 + se(n, na.rm = na.rm)^ 2)
}

# Calculate variance of deviance as SED
deviance$se <-
  fluxes %>% 
  select(treat, passage, estimate) %>% 
  mutate(x = 1:nrow(fluxes)) %>% 
  pivot_wider(names_from = treat, values_from = estimate) %>% 
  group_by(passage) %>% 
  summarize(sed = sed(p, n, na.rm = TRUE), .groups = "drop") %>%
  pull(sed)

# Fit deviance model
fit <- rma(deviance ~ passage, se^2, method="FE", data = deviance)

# Plot deviance
ggplot(deviance, aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
  geom_pointrange(color = 'darkorange2') + 
  geom_abline(intercept = fit$beta[1], slope = fit$beta[2])

# Take deviance model parameters and export into a data.frame
rma_output <- data.frame(
Estimate = fit$beta,
se = fit$se,
z.value = fit$zval,
p.value = fit$pval,
upper.ci = fit$ci.lb,
lower.ci = fit$ci.ub
)

# Fit deviance model without passage 2
fit_no2 <- 
filter(deviance, passage != 2) %>% 
rma(deviance ~ passage, se^2, method="FE", data = .)

# Take deviance model withou passage 2 parameters and export into a data.frame

rma_output_no2 <- data.frame(
Estimate = fit$beta,
se = fit$se,
z.value = fit$zval,
p.value = fit$pval,
upper.ci = fit$ci.lb,
lower.ci = fit$ci.ub
)

# R (response to selection) for the no2 model
R_no2 <- rma_output_no2[[2, 1]]
R <- rma_output[[2, 1]]

# Plot the no2 model
filter(deviance, passage != 2) %>% 
ggplot(., aes(x = passage, y = deviance, ymin = deviance - se, ymax = deviance + se)) + 
  geom_pointrange(color = 'darkorange2') + 
  geom_abline(intercept = fit_no2$beta[1], slope = fit_no2$beta[2])

# Print the full model
rma_output %>% 
  kable() %>% 
  kable_styling()

write.csv(deviance, '../Output/deviance.tsv')
```

@@

```{r create_heritability_data, message=FALSE, warning=FALSE}
selected_jars <- read_csv('../Data/selected.csv')

# Pull out the selected parental jars
selected <- 
  fluxes[paste0(fluxes$passage, fluxes$jar) %in% paste0(selected_jars$passage, selected_jars$jar), ] %>%
  select(treat, estimate, passage)

# Pull out all parental jars
parental <-
  fluxes %>% 
  select(passage, treat, estimate)

# Pull out all offspring jars
offspring <- 
  fluxes %>%
  select(treat, estimate, passage) %>% 
  filter(passage != 1) %>% 
  mutate(passage = passage - 1)
```

```{r heritability_mean}

# Calculate means for east generation
parental_mean <-
  parental %>% 
  group_by(passage, treat) %>% 
  summarize(parental = mean(estimate), .groups = "drop")

offspring_mean <-
  offspring %>% 
  group_by(passage, treat) %>% 
  summarize(offspring = mean(estimate), .groups = "drop")

selected_mean <-
  selected %>% 
  group_by(passage, treat) %>% 
  summarize(selected = mean(estimate), .groups = "drop")

# Combine all parent/offspring data into one data frame
heritability <- 
  parental_mean %>% 
  left_join(offspring_mean, by = c("passage", "treat")) %>% 
  left_join(selected_mean, by = c("passage", "treat")) %>% 
  #filter(treat == "p") %>% 
  filter(passage != 5) %>% # Remove passage 5, which has no offspring
  ungroup()

@@

``` {r generation_h2}

######################################################################
# This is where you stopped
# Figure out why h2 is different from h2_calc_method
# Finish writing your email to brendan about heritability
######################################################################

# Function for the Breeder's equation

breeders <- function(R, S) {
  R / S
}

# Calculate h2 (heritability) using math, not regression
# For each generation and then calculate the mean/se
h2_calc_method <-
  heritability %>%
  filter(treat == "p") %>%
  mutate(S = selected - parental) %>%
  mutate(R = offspring - parental) %>%
  mutate(h2 = breeders(R, S)) %>%
  summarize(S = mean(S),
            R = mean(R),
            se = se(h2),
            h2 = mean(h2))

h2_calc_method_no_2 <-
  heritability %>%
  filter(treat == "p") %>%  
  filter(passage != 2) %>%
  mutate(S = selected - parental) %>%
  mutate(R = offspring - parental) %>%
  mutate(h2 = breeders(R, S)) %>%
  summarize(S = mean(S),
            R = mean(R),
            se = se(h2),
            h2 = mean(h2))
h2 <- 
  h2_calc_method %>% 
  mutate(method = "calc p")
h2_no2 <- 
  h2_calc_method_no_2 %>% 
  mutate(method = "calc p no 2")

h2 <- rbind(h2, h2_no2)
```

```{r h2_slope_method}

# Calculate S (selection differential) with/without passage 2

S <-
  heritability %>%
  filter(treat == "p") %>%
  mutate(S = abs(selected - parental)) %>%
  summarize(S = mean(S)) %>% 
  pull()
S_no2 <-
  heritability %>%
  filter(treat == "p") %>%
  filter(passage != 2) %>% 
  mutate(S = abs(selected - parental)) %>%
  summarize(S = mean(S)) %>% 
  pull()

# For the equivalent plot of Falconer and MacKay
# Scale and center all fluxes
heritability_scaled <-
  heritability %>% 
  pivot_longer(parental:selected) %>% 
  mutate(value = scale(value)) %>% 
  pivot_wider(names_from = "name", values_from = "value")

# Calculate R and S 
R_and_S <-
  heritability_scaled %>% 
  filter(treat == "p") %>% 
  summarize(R = mean(offspring),
            S = mean(selected))

# Calculate scaled h2
R <- R_and_S[[1, 'R']]
S <- R_and_S[[1, 'S']]
h2 <- breeders(R, S)

# Plot scaled h2
ggplot(data = heritability_scaled, aes(x = selected, y = offspring)) +
  geom_point(aes(shape = treat)) +
  scale_shape_manual(values = c(1,19), name = "Treatment", labels = c('Neutral', 'Positive')) +
  stat_smooth(se = FALSE, method = "lm", formula = y ~ x, color = 'black') +
  labs(x = "Selected Parents", y = "Offspring") +
  geom_segment(aes(x = 0, y = R, xend = S, yend = R), linetype = 3) +
  geom_segment(aes(x = S, y = 0, xend = S, yend = R), linetype = 3) +
  geom_hline(yintercept = 0) +
  geom_vline(xintercept = 0) +
  geom_text(label = "R", x = -0.1, y = R) +
    geom_text(label = "S", x = S, y = -0.2)

```{r}
heritability %>% 
  ggplot(aes(parental, offspring, color = treat)) +
  geom_point() +
  labs(x = "Parental", y = "Offspring") +
  theme_bw() +
  scale_color_discrete(name = "Treatment", labels = c('Neutral', 'Positive'))

summary(lm(offspring ~ parental, data = heritability))
```
```{r}
heritability %>% 
  ggplot(aes(selected, offspring, color = treat)) +
  geom_point() +
  labs(x = "Selected", y = "Offspring") +
  theme_bw() +
  scale_color_discrete(name = "Treatment", labels = c('Neutral', 'Positive'))
```
